Overview

### 2023-12-07 analysis/001_overview.Rmd
Fisrt draft country plot and save figures in dedicated folder

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

@@

```{r, country_plot, fig.asp=0.8}
ggplot(top_20_plot_df, aes(
    gsub("^20", "'", as.character(cr_year)),
    value / all_total,
    group = type,
    fill = type
  )) +
  geom_area(stat = "identity") +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  facet_wrap( ~ country_name_fct,
              nrow = 3,
              labeller = as_labeller(country_facet_labels)) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
  labs(y = "OA Share", x = "2018 - 2022") +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    strip.text = ggtext::element_markdown(size = 6),
    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
    panel.spacing = unit(1, "lines")
  )
```

>>>>

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300,
  dev = c("png", "pdf"),
  fig.path = "fig/"
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

@@

```{r, country_plot, fig.asp=0.8}
ggplot(top_20_plot_df, aes(
    gsub("^20", "'", as.character(cr_year)),
    value / all_total,
    group = type,
    fill = type
  )) +
  geom_area(stat = "identity") +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  facet_wrap( ~ country_name_fct,
              nrow = 3,
              labeller = as_labeller(country_facet_labels)) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = "2018 - 2022") +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    strip.text = ggtext::element_markdown(size = 6),
    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
    panel.spacing = unit(1, "lines")
  )
```

here a scatterplot and a classifciation of oecd / brics / other. first, the scatterplot 2022!

```{r}
oecd <- readr::read_csv(here::here("data", "oecd.csv")) |>
  mutate(country_code = countrycode::countrycode(iso3, "iso3c", "iso2c"))

country_scatter_df <- country_df |>
#  filter(cr_year == "2022", !is.na(country_code)) |>
  mutate(cat = case_when(
    country_code %in% oecd$country_code ~ "OECD",
    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
    .default = "Other"
  ))

country_scatter_22 <- country_scatter_df |>
  filter(cr_year == "2022", !is.na(country_code)) |>
  ggplot(aes(all, oa)) +
  geom_point(aes(color = cat, size = ta_oa), alpha = 0.8) +
  scale_color_manual(NULL, values = c("OECD" = "#3fc1c9", "BRICS" = "#fc5185", "Other" = "#e1e1e1")) +
  geom_smooth(aes(all, oa), method=lm) +
  scale_size("TA OA", guide = "none") +
  ggrepel::geom_text_repel(
    data = filter(country_scatter_df, cr_year == 2022, cat %in% c("OECD", "BRICS"), all > 50000),
    aes(x = all, y = oa, label = country_code),
    force = 50,
    box.padding = 0.5,
    point.padding = 0.8,
    size = 3,
    max.iter = 1000000,
    max.overlaps = Inf,
    min.segment.length = 0,
    arrow = arrow(length = unit(0.010, "npc"))) +
  scale_x_continuous(labels =  scales::number_format(big.mark = ","), guide = guide_axis(angle = 45)) +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Total articles", y = "OA articles in hybrid journals") +
  theme(legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal")

country_bar <- country_scatter_df |>
  group_by(cr_year, cat) |>
  summarise(all = sum(all), oa = sum(oa), ta_oa = sum(ta_oa)) |>
  mutate(closed = all - oa) |>
  pivot_longer(cols = c(oa, ta_oa, closed), names_to = "oa_type") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
  geom_area() +
  facet_wrap(~ cat) +
  scale_fill_manual(NULL, values = c(oa = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e98b56"),
    labels = c(oa = "OA", ta_oa = "Agreement", closed = "Closed")) +
  theme_minimal() +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Publication year", y = "Total articles") +
  theme(legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal")
```

```{r country_patch}
library(patchwork)
wrap_plots(
  A = country_scatter_22,
  B = country_bar,
  design = "AB",
  widths = c(0.35, 0.65)
) &
  theme(
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```
